Minimal trap spaces of Boolean networks are maximal siphons of their Petri net encoding

### 2022-12-22 manuscripts/TCS2022/tcs2022.tex
Updates from Overleaf

\setlength{\tabcolsep}{4pt}
\begin{table}[!htb]
  \caption{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
    Column \(n\) denotes the number of nodes of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}
  \centering%
  \label{tab:pyboolnet_repo}
  \begin{tabular}{rlrrrrrr}
    \toprule
    & model & $n$ & $|M|$ & LQM & PBN & mpbn & Trappist \\ \midrule
    \rownb & arellano\_rootstem & 9 & 4 & - & 0.05 & 0.01 & \textbf{0.20}\\
    \rownb & calzone\_cellfate & 28 & 27 & - & 0.03 & \textbf{NM} & 0.03\\
    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & - & 0.05 & 0.02 & 0.03\\
    \rownb & davidich\_yeast & 10 & 12 & - & 0.04 & 0.01 & 0.02\\
    \rownb & dinwoodie\_life & 15 & 7 & - & 0.03 & 0.01 & 0.01\\
    \rownb & dinwoodie\_stomatal & 13 & 1 & - & 0.03 & 0.01 & 0.01\\
    \rownb & faure\_cellcycle & 10 & 2 & - & 0.04 & 0.01 & 0.02\\
    \rownb & grieco\_mapk & 53 & 18 & - & 0.04 & 0.02 & 0.02\\
    \rownb & irons\_yeast & 18 & 1 & - & 0.05 & 0.01 & 0.02\\
    \rownb & jaoude\_thdiff & 103 & $>1000$ & - & \textbf{1.44} & \textbf{0.90} & 0.09\\
    \rownb & klamt\_tcr & 40 & 8 & - & 0.04 & 0.01 & 0.03\\
    \rownb & krumsiek\_myeloid & 11 & 6 & - & 0.03 & 0.01 & 0.01\\
    \rownb & multivalued & 13 & 4 & - & 0.03 & 0.01 & 0.01\\
    \rownb & n12c5 & 11 & 5 & - & \textbf{35.16} & 0.01 & 0.02\\
    \rownb & n3s1c1a & 2 & 2 & - & 0.02 & 0.01 & 0.01\\
    \rownb & n3s1c1b & 2 & 2 & - & 0.02 & 0.01 & 0.01\\
    \rownb & n5s3 & 4 & 3 & - & 0.03 & \textbf{NM} & 0.01\\
    \rownb & n6s1c2 & 5 & 3 & - & 0.03 & 0.01 & 0.01\\
    \rownb & n7s3 & 6 & 3 & - & 0.02 & 0.01 & 0.01\\
    \rownb & raf & 3 & 2 & - & 0.02 & 0.01 & 0.01\\
    \rownb & randomnet\_n15k3 & 15 & 3 & - & 0.03 & \textbf{NM} & 0.01\\
    \rownb & randomnet\_n7k3 & 7 & 10 & - & 0.03 & \textbf{NM} & 0.01\\
    \rownb & remy\_tumorigenesis & 34 & 25 & - & \textbf{2.14} & 0.02 & 0.02\\
    \rownb & saadatpour\_guardcell & 13 & 1 & - & 0.03 & 0.01 & 0.01\\
    \rownb & selvaggio\_emt & 56 & $>1000$ & - & \textbf{1.02} & \textbf{0.52} & 0.09\\
    \rownb & tournier\_apoptosis & 12 & 3 & - & 0.04 & 0.01 & 0.01\\
    \rownb & xiao\_wnt5a & 7 & 4 & - & 0.03 & 0.01 & 0.01\\
    \rownb & zhang\_tlgl & 60 & 156 & - & 0.22 & \textbf{NM} & 0.05\\
    \rownb & zhang\_tlgl\_v2 & 60 & 258 & - & 0.09 & 0.15 & 0.02\\
    \bottomrule
  \end{tabular}
\end{table}
\setcounter{rownumber}{0}

@@

\setlength{\tabcolsep}{4pt}
\begin{table}[!h]
  \caption{Results on the real-world models from the BBM repository.}
  \centering%
  \label{tab:bbm_repo}
  \begin{tabular}{l||r|r|r||r}
    \toprule
    Method & 1 min & 2 min & 3 min & Time average (s) \\ \midrule
    PyBoolNet & - & - & - & 0.05 \\
    mpbn & - & - & - & 0.05 \\
    Trappist & - & - & - & 0.05 \\
    \bottomrule
  \end{tabular}
\end{table}
\setcounter{rownumber}{0}

@@

\begin{table}[!htb]
  \caption{Timing comparisons between PyBoolNet (PBN), mpbn and Trappist on selected models from the literature.
    Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
    ``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes. ``NM'' indicates a non-locally-monotonic model.}
  \centering%
  \label{tab:result_real}
  \begin{tabular}{rlrrrrrrr}
    \toprule
    & & & & & PBN & mpbn & \multicolumn{2}{c}{Trappist} \\
    \cmidrule(lr){8-9}
    & model & \(n\) & \(s\) & \(|M|\) & 1000 & 1000 & 1000 & all \\
    \midrule
    \rownb & inflammatory-bowel~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 47 & 0 & 1 & DNF & NM & 1.26 & 0.69\\
    \rownb & T-LGL-survival~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 61 & 7 & 318 & 0.84 & NM & 0.01 & 0.01\\
    \rownb & butanol-production~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 66 & 13 & 8192 & 0.71 & NM & 0.01 & 0.02\\
    \rownb & colon-cancer~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 70 & 1 & 10 & 0.20 & NM & 0.02 & 0.01\\
    \rownb & mast-cell-activation~\cite{aghamiri2020automated} & 73 & 19 & $>1000$ & 0.76 & NM & 0.01 & DNF\\
    \rownb & IL-6-signalling~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 86 & 15 & 32768 & 1.29 & NM & 0.01 & 0.01\\
    \rownb & Corral-ThIL-17-diff~\cite{corral2021interplay} & 92 & 16 & $>1000$ & DNF & NM & 0.03 & DNF\\
    \rownb & Korkut-2015~\cite{lee2019signal} & 99 & 12 & 18556 & DNF & 1.04 & 0.06 & 0.08\\ \midrule
    \rownb & adhesion-cip-migration~\cite{guberman2020boolean} & 121 & 4 & 78 & 36.80 & 0.35 & 0.27 & 0.09\\
    \rownb & interferon-1~\cite{ostaszewski2021covid19} & 121 & 55 & $>1000$ & 10.11 & NM & 0.02 & DNF\\
    \rownb & TCR-TLR5-signaling~\cite{rodriguez2019cooperation} & 130 & 5 & 48 & 2.06 & NM & 0.03 & 0.02\\
    \rownb & influenza-replication~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 131 & 11 & 10128 & 46.67 & NM & 0.02 & 0.02\\
    \rownb & prostate-cancer~\cite{Montagud2021} & 133 & 11 & 2760 & DNF & NM & 0.16 & 0.04\\
    \rownb & HIV-1~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 138 & 14 & 39424 & DNF & NM & 0.06 & 0.06\\
    \rownb & fibroblasts~\cite{helikar2008emergent} & 139 & 9 & $>1000$ & DNF & NM & 0.08 & DNF\\
    \rownb & HMOX-1-pathway~\cite{ostaszewski2021covid19} & 145 & 56 & $>1000$ & 6.52 & NM & 0.02 & DNF\\
    \rownb & kynurenine-pathway~\cite{ostaszewski2021covid19} & 150 & 72 & $>1000$ & DNF & NM & 0.19 & DNF\\
    \rownb & virus-replication-cycle~\cite{ostaszewski2021covid19} & 154 & 25 & $>1000$ & DNF & NM & 0.03 & DNF\\
    \rownb & immune-system~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 164 & 13 & $>1000$ & DNF & NM & 0.08 & DNF\\
    \rownb & RA-apoptosis~\cite{aghamiri2020automated} & 180 & 59 & $>1000$ & DNF & NM & 0.02 & DNF\\
    \rownb & MAPK~\cite{aghamiri2020automated} & 181 & 37 & $>1000$ & 87.13 & NM & 0.03 & DNF\\
    \rownb & er-stress~\cite{ostaszewski2021covid19} & 182 & 75 & $>1000$ & 17.88 & NM & 0.03 & DNF\\
    \rownb & cascade-3~\cite{Tsirvouli2020} & 183 & 0 & 1 & 101.18 & NM & 0.40 & 0.07\\
    \rownb & CHO-2016~\cite{lee2019signal} & 200 & 13 & 13312 & DNF & 2.45 & 0.07 & 0.08\\
    \rownb & T-cell-check-point~\cite{hernandez2020computational} & 218 & 14 & $>1000$ & 69.94 & NM & 0.05 & DNF\\
    \rownb & ErbB-receptor-signaling~\cite{helikar2013comprehensive} & 247 & 22 & $>1000$ & DNF & NM & 0.24 & DNF\\
    \rownb & macrophage-activation~\cite{DBLP:journals/bmcsb/HelikarKMBRMWSLR12} & 321 & 19 & $>1000$ & 16.43 & NM & 0.05 & DNF\\
    \rownb & cholocystokinin~\cite{aghamiri2020automated} & 383 & 74 & $>1000$ & 1.42 & NM & 0.08 & DNF\\
    \rownb & Alzheimer~\cite{aghamiri2020automated} & 762 & 237 & $>1000$ & DNF & NM & 0.29 & DNF\\ \midrule
    \rownb & KEGG-network~\cite{DBLP:journals/bmcsb/Kwon16} & 1659 & 521 & $>1000$ & DNF & 20.76 & 3.40 & DNF\\
    \rownb & human-network~\cite{kim2011reduction} & 1953 & 669 & $>1000$ & DNF & 23.62 & 5.93 & DNF\\
    \rownb & SN-5~\cite{kim2013rmod} & 2746 & 829 & $>1000$ & DNF & 28.94 & DNF & DNF\\
    \rownb & turei-2016~\cite{lee2019signal} & 4691 & 1257 & ??? & DNF & DNF & DNF & DNF\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[!htb]
  \caption{Results on N-K models.}
  \centering%
  \label{tab:result_N_K}
  \begin{tabular}{rrrrrrr}
    \toprule
    & \multicolumn{2}{c}{1 min} & \multicolumn{2}{c}{2 min} & \multicolumn{2}{c}{3 min} \\
    \cmidrule(lr){2-7}
    \(n\) & PBN & Trappist & PBN & Trappist & PBN & Trappist \\
    \midrule
    100 & - & - & - & - & - & - \\
    150 & - & - & - & - & - & - \\
    200 & - & - & - & - & - & - \\
    250 & - & - & - & - & - & - \\
    300 & - & - & - & - & - & - \\
    350 & - & - & - & - & - & - \\
    400 & - & - & - & - & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[!htb]
  \caption{Results on scale-free models.}
  \centering%
  \label{tab:result_N_K}
  \begin{tabular}{rrrrrrr}
    \toprule
    & \multicolumn{2}{c}{1 min} & \multicolumn{2}{c}{2 min} & \multicolumn{2}{c}{3 min} \\
    \cmidrule(lr){2-7}
    \(n\) & PBN & Trappist & PBN & Trappist & PBN & Trappist \\
    \midrule
    500 & - & - & - & - & - & - \\
    750 & - & - & - & - & - & - \\
    1000 & - & - & - & - & - & - \\
    1250 & - & - & - & - & - & - \\
    1500 & - & - & - & - & - & - \\
    1750 & - & - & - & - & - & - \\
    2000 & - & - & - & - & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

\setlength{\tabcolsep}{4pt}
\begin{table}[!htb]
  \caption{Timing comparisons between PyBoolNet, mpbn and Trappist on the PyBoolNet repository.
    Column \(n\) denotes the number of nodes of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds, asking only for the first 1000 trap spaces.
    A number in bold indicates a ratio greater than three compared to the best result. ``NM'' indicates a non-locally-monotonic model.}
  \centering%
  \label{tab:pyboolnet_repo}
  \begin{tabular}{rlrrrrrr}
    \toprule
    & model & $n$ & $|M|$ & LQM & PBN & mpbn & Trappist \\ \midrule
    \rownb & arellano\_rootstem & 9 & 4 & 0.13 & 0.05 & 0.01 & \textbf{0.20}\\
    \rownb & calzone\_cellfate & 28 & 27 & 0.12 & 0.03 & \textbf{NM} & 0.03\\
    \rownb & dahlhaus\_neuroplastoma & 23 & 32 & 0.11 & 0.05 & 0.02 & 0.03\\
    \rownb & davidich\_yeast & 10 & 12 & 0.11 & 0.04 & 0.01 & 0.02\\
    \rownb & dinwoodie\_life & 15 & 7 & 0.11 & 0.03 & 0.01 & 0.01\\
    \rownb & dinwoodie\_stomatal & 13 & 1 & 0.10 & 0.03 & 0.01 & 0.01\\
    \rownb & faure\_cellcycle & 10 & 2 & 0.11 & 0.04 & 0.01 & 0.02\\
    \rownb & grieco\_mapk & 53 & 18 & 0.19 & 0.04 & 0.02 & 0.02\\
    \rownb & irons\_yeast & 18 & 1 & 0.12 & 0.05 & 0.01 & 0.02\\
    \rownb & jaoude\_thdiff & 103 & $>1000$ & N/A & \textbf{1.44} & \textbf{0.90} & 0.09\\
    \rownb & klamt\_tcr & 40 & 8 & 0.11 & 0.04 & 0.01 & 0.03\\
    \rownb & krumsiek\_myeloid & 11 & 6 & 0.10 & 0.03 & 0.01 & 0.01\\
    \rownb & multivalued & 13 & 4 & 0.10 & 0.03 & 0.01 & 0.01\\
    \rownb & n12c5 & 11 & 5 & 0.11 & \textbf{35.16} & 0.01 & 0.02\\
    \rownb & n3s1c1a & 2 & 2 & 0.10 & 0.02 & 0.01 & 0.01\\
    \rownb & n3s1c1b & 2 & 2 & 0.09 & 0.02 & 0.01 & 0.01\\
    \rownb & n5s3 & 4 & 3 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
    \rownb & n6s1c2 & 5 & 3 & 0.10 & 0.03 & 0.01 & 0.01\\
    \rownb & n7s3 & 6 & 3 & 0.11 & 0.02 & 0.01 & 0.01\\
    \rownb & raf & 3 & 2 & 0.10 & 0.02 & 0.01 & 0.01\\
    \rownb & randomnet\_n15k3 & 15 & 3 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
    \rownb & randomnet\_n7k3 & 7 & 10 & 0.10 & 0.03 & \textbf{NM} & 0.01\\
    \rownb & remy\_tumorigenesis & 34 & 25 & 0.15 & \textbf{2.14} & 0.02 & 0.02\\
    \rownb & saadatpour\_guardcell & 13 & 1 & 0.10 & 0.03 & 0.01 & 0.01\\
    \rownb & selvaggio\_emt & 56 & $>1000$ & N/A & \textbf{1.02} & \textbf{0.52} & 0.09\\
    \rownb & tournier\_apoptosis & 12 & 3 & 0.10 & 0.04 & 0.01 & 0.01\\
    \rownb & xiao\_wnt5a & 7 & 4 & 0.10 & 0.03 & 0.01 & 0.01\\
    \rownb & zhang\_tlgl & 60 & 156 & 0.60 & 0.22 & \textbf{NM} & 0.05\\
    \rownb & zhang\_tlgl\_v2 & 60 & 258 & 0.64 & 0.09 & 0.15 & 0.02\\
    \bottomrule
  \end{tabular}
\end{table}
\setcounter{rownumber}{0}

@@

\setlength{\tabcolsep}{4pt}
\begin{table}[!h]
  \caption{Results on the real-world models from the BBM repository.}
  \centering%
  \label{tab:bbm_repo}
  \begin{tabular}{l||r|r|r||r}
    \toprule
    Method & 1 min & 2 min & 3 min & Time average (s) \\ \midrule
    PyBoolNet & - & - & - & 0.05 \\
    mpbn & - & - & - & 0.05 \\
    Trappist-maxSAT & - & - & - & 0.05 \\
    Trappist-CLP & - & - & - & 0.05 \\
    Trappist-ILP & - & - & - & 0.05 \\
    Trappist-ASP & - & - & - & 0.05 \\
    \bottomrule
  \end{tabular}
\end{table}
\setcounter{rownumber}{0}

@@

Column \(n\) (resp.\ \(s\)) denotes the number of nodes (resp.\ source nodes) of each model. Column \(|M|\) denotes the number of minimal trap spaces and for each method is given the computation time in seconds.
``DNF'' means that the method did not finish the computation (stopping at the first 1000 minimal trap spaces, or all for the last column) within the timeout of two minutes. ``NM'' indicates a non-locally-monotonic model.

\begin{table}[!htb]
  \caption{Timing comparisons between bioLQM (LQM), PyBoolNet (PBN), mpbn and the four variants of Trappist on selected models from the literature.}
  \centering%
  \label{tab:result_real}
  \begin{tabular}{rlrrrrrrrrr}
    \toprule
    & & & & & & & \multicolumn{4}{c}{Trappist}\\
    \cmidrule(rr){8-11}
    & model & \(n\) & \(|M|\) & LQM & PBN & mpbn & SAT & CLP & ILP & ASP \\
    \midrule % n < 100
    \rownb & Arabidopsis\_thaliana & 15 & 8 & 0.15 & - & - & - & - & - & - \\
    \rownb & Bcell & 73 & 72 & 0.21 & - & - & - & - & - & - \\
    \rownb & Cacace\_TdevModel & 61 & 28 & 1.21 & - & NM & - & - & - & - \\
    \rownb & Corral\_ThIL17diff & 92 & - & N/A & - & - & - & - & - & - \\
    \rownb & DNA\_damage & 26 & - & 0.23 & - & NM & - & - & - & - \\
    \rownb & Drosophila & 52 & - & 0.33 & - & - & - & - & - & - \\
    \rownb & EMT & 69 & - & - & - & - & - & - & - & - \\
    \rownb & mast\_cell & 73 & - & N/A & - & - & - & - & - & - \\
    \rownb & FT-GRN & 23 & - & DNF & - & NM & - & - & - & - \\
    \rownb & hedgehog & 65 & - & N/A & - & - & - & - & - & - \\
    \rownb & metastatic & 10 & - & 0.11 & - & NM & - & - & - & - \\
    \rownb & Pancreatic\_Cancer & 43 & - & N/A & - & - & - & - & - & - \\
    \rownb & Pluripotent & 36 & - & 0.37 & - & NM & - & - & - & - \\
    \rownb & Rho-GTPases & 33 & 1 & 0.21 & - & - & - & - & - & - \\
    \rownb & Pluripotency & 36 & - & DNF & - & NM & - & - & - & - \\
    \rownb & p53\_high\_dna & 16 & 1 & 0.40 & - & NM & - & - & - & - \\
    \rownb & p53\_low\_dna & 16 & 1 & 0.38 & - & NM & - & - & - & - \\

    \midrule % 100 <= n < 200
    \rownb & angiofull & 142 & - & N/A & - & - & - & - & - & - \\
    \rownb & COVID\_19 & 111 & - & N/A & - & - & - & - & - & - \\
    \rownb & EMT\_Mech & 136 & - & DNF & - & - & - & - & - & - \\
    \rownb & EMT\_Mech\_TGFbeta & 150 & - & DNF & - & - & - & - & - & - \\
    \rownb & MAPK & 181 & - & N/A & - & - & - & - & - & - \\
    \rownb & macrophage & 136 & - & N/A & - & - & - & - & - & - \\
    \rownb & RA\_apoptosis & 180 & - & N/A & - & - & - & - & - & - \\
    \rownb & Adhesion\_CIP & 121 & - & 57.33 & - & - & - & - & - & - \\
    \rownb & angiogenesis & 141 & - & N/A & - & - & - & - & - & - \\

    \midrule % 200 <= n
    \rownb & T-cell-co-receptor & 206 & - & N/A & - & - & - & - & - & - \\
    \rownb & Snf1-pathway & 202 & - & N/A & - & - & - & - & - & - \\
    \rownb & Mycobacterium & 317 & - & N/A & - & - & - & - & - & - \\
    \rownb & Leishmania & 342 & - & N/A & - & - & - & - & - & - \\
    \rownb & TcellCheckPoint & 218 & - & N/A & - & NM & - & - & - & - \\
    \rownb & Cholocystokinin & 383 & - & N/A & - & - & - & - & - & - \\
    \rownb & Alzheimer & 762 & - & N/A & - & NM & - & - & - & - \\

    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[!htb]
  \caption{Results on N-K models.}
  \centering%
  \label{tab:result_N_K}
  \begin{tabular}{rrrrrrrr}
    \toprule
    & & & & \multicolumn{4}{c}{Trappist}\\
    \cmidrule(rr){5-8}
    \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
    \midrule
    100 & - & - & - & - & - & - & - \\
    150 & - & - & - & - & - & - & - \\
    200 & - & - & - & - & - & - & - \\
    250 & - & - & - & - & - & - & - \\
    300 & - & - & - & - & - & - & - \\
    350 & - & - & - & - & - & - & - \\
    400 & - & - & - & - & - & - & - \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[!htb]
  \caption{Results on scale-free models.}
  \centering%
  \label{tab:result_scale_free}
  \begin{tabular}{rrrrrrrr}
    \toprule
    & & & & \multicolumn{4}{c}{Trappist}\\
    \cmidrule(rr){5-8}
    \(n\) & LQM & mpbn & PBN & maxSAT & CLP & ILP & ASP \\
    \midrule
    500 & - & - & - & - & - & - & - \\
    750 & - & - & - & - & - & - & - \\
    1000 & - & - & - & - & - & - & - \\
    1250 & - & - & - & - & - & - & - \\
    1500 & - & - & - & - & - & - & - \\
    1750 & - & - & - & - & - & - & - \\
    2000 & - & - & - & - & - & - & - \\
    \bottomrule
  \end{tabular}
\end{table}
